---
title: "microbe_metabolite_correlation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spearman correlation

this is a demo of spearman correlation, used as a proxy to rare co-presence/absence of metabolites/microbes. The code is not optimized and we used a computer with a large amount of RAM memory (64GB). 

Load data and label rows with sample names:

```{r eval=FALSE}
tab1 <- read.csv("20160913_AMG_optimus_1M.csv")
met1 <- read.delim("160803_AMG_meta.txt")
tab1 <- merge(met1, tab1, by.x = "X.SampleID", by.y = "X.SampleID")
tab2 <- read.delim("table.from_biom_w_taxonomy_embriete.txt", skip=1)
tab2tmp <- t(tab2[,-c(1,109)]) 
colnames(tab2tmp) <- tab2[,109] 
tab2 <- tab2tmp 
```

Normalize the samples by total count and total ion count for microbe/metabolites and make sure the samples are in the same order:

```{r eval=FALSE}
tab2c <- t(apply(tab2, 1, function(x) x/sum(x))) 
tab1c <- t(apply(tab1[,-c(1:64)], 1, function(x) x/sum(x))) 
rownames(tab1c) <- as.character(tab1[,1]) 
rownames(tab2c) <- gsub("0+$", "", rownames(tab2c)) 
rownames(tab2c) <- gsub("^X", "", rownames(tab2c)) 
```

Perform correlation analysis:

```{r eval=FALSE}
flattenCorrMatrix <- function(cormat, pmat) {
   ut <- upper.tri(cormat)
   data.frame(
     row = rownames(cormat)[row(cormat)[ut]],
     column = rownames(cormat)[col(cormat)[ut]],
     cor  =(cormat)[ut],
     p = pmat[ut]
     )
 }
 
library(Hmisc)
 
res<-rcorr(as.matrix(ctab), type="spearman")

flat <- flattenCorrMatrix(res$r, res$P)
flat <- flat[(flat[,1] %in% colnames(tab1c)) & (flat[,2] %in% colnames(tab2c)),]

flat <-   flat[!is.na(flat[,4]),]
# The p-value is biased by ties in this sparse data
#p.adj <- p.adjust(flat[,4], method="fdr")
#flat <- cbind(flat[,1:4], p.adj)
#sel_flat <- flat[flat[,5] < 0.05,]
```

Record the number of samples in which you have counts/intensity different of zero. 

```{r eval=FALSE}
nozero_met <- sapply(unique(as.character(sel_flat[,1])), function(x) sum(tab1c[, x]!=0))
nozero_micro <- sapply(unique(as.character(sel_flat[,2])), function(x) sum(tab2c[, x]!=0))

sel_flat <- cbind(sel_flat, 0, 0)
sel_flat[,6] <- nozero_met[as.character(sel_flat[,1])]
sel_flat[,7] <- nozero_micro[as.character(sel_flat[,2])]

colnames(sel_flat)[6:7] <- c("nozero_met", "nozero_micro")
head(sel_flat)
write.csv(sel_flat, "all_spearman_corr.csv", row.names=FALSE)

```


## Generating cytoscape correlation network

Here we use Cytoscape 2.8 and its plugin CytoscapeRPC to automatically output the network from R to Cytoscape. Install the plugin, open Cytoscape and Activate CytoscapeRPC before running the R code.

Load the correlations calculated in the previous step and sub-select a number of samples where we have signal of microbes and metabolites, to subsequently filter by the desired correlation threshold and create a graph.

```{r eval=FALSE}
ctable <- read.csv("all_spearman_corr.csv")
sum(ctable[,"nozero_met"] > 1 & ctable[,"nozero_met"] < 10 & ctable[,"nozero_micro"] > 1 & ctable[,"nozero_micro"] < 10)
net1 <- ctable[,"nozero_met"] > 1 & ctable[,"nozero_met"] < 10 & ctable[,"nozero_micro"] > 1 & ctable[,"nozero_micro"] < 10

library(RCytoscape)

ctable2 <- ctable[net1,][ctable[net1,3]>0.72,]

ctable2 <- ctable2[-which(duplicated(ctable2[,1:2])),]
gr <- ftM2graphNEL(as.matrix(ctable2[,1:2]))
```

Format the graph and export to Cytoscape.

```{r eval=FALSE}
g2 <- new("graphNEL", edgemode = "undirected")
  i <- 1
  system.time(lapply(gr@edgeL, function(x) {
  if (length(x$edges)) {
  gr@edgeL[[i]]$edges <<- x$edges[x$edges > i]
  }
  i <<- i + 1
}))
cw5 <- new.CytoscapeWindow("corWindow", graph = g2, overwriteWindow = TRUE)
cw5@graph <- gr
layoutNetwork(cw5, layout.name = "jgraph-spring")
displayGraph(cw5)
setNodeShapeDirect(cw5, as.character(ctable2[,2]), "hexagon")
setNodeColorDirect(cw5, as.character(ctable2[,2]), "#00CD66")
redraw(cw5)
```

Pair GNPS's precursor ions to MS1 ions and export the information to Cytoscape.

```{r eval=FALSE}
gnps <- read.delim("ProteoSAFe-METABOLOMICS-SNETS-25fa4086-download_clustered_spectra/clusterinfosummarygroup_attributes_withIDs/558c9adb70e441bdbdf646af9ca2c015..out")

match.peaks <- function(query, ref, mz.tol = 0.2, rt.tol = 30) {
  mzdiff <- abs(ref[,1]-query[1])
  rtdiff <- abs(ref[,2]-query[2])
  mtch <- which((mzdiff <= mz.tol) & (rtdiff <= rt.tol))
  mtch
}

mv <- sapply(sub("^X", "", as.character(ctable2[,1])), function(x) paste(strsplit(x, "\\.")[[1]][1:2], collapse="."))
rtv <- sapply(sub("^X", "", as.character(ctable2[,1])), function(x) paste(strsplit(x, "\\.")[[1]][3]))

fetm <- cbind(mv, rtv)
fetm <- apply(fetm, 2, as.numeric)

tabgnps <- gnps[,c("precursor.mass", "RTMean")]
mt <- apply(fetm, 1, match.peaks, tabgnps)

LibraryID <- lapply(mt, function(x) if(length(x)) paste(gnps[x, "LibraryID"], collapse = ";") else "")
LibraryID <- unlist(LibraryID)
ProteoSAFeClusterLink <- lapply(mt, function(x) if(length(x)) paste(gnps[x, "ProteoSAFeClusterLink"], collapse = ";") else "")
ProteoSAFeClusterLink <- unlist(ProteoSAFeClusterLink)
attr.mt <- data.frame(Name=ctable2[,1], mzmed=fetm[,1], rtmed=fetm[,2], LibraryID, ProteoSAFeClusterLink)
for(i in 2:ncol(attr.mt)) setNodeAttributesDirect(cw5, colnames(attr.mt)[i], "String", as.character(attr.mt[,1]), as.character(attr.mt[,i]))
redraw(cw5)
```
